---
title: "Penguins ANOVA"
author: "Nathaniel Burola"
date: "November 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read-in penguins.csv data 

```{r, include = FALSE}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read.csv("penguins.csv")
```

Our question: 

1. Is there a significant difference in mean flipper length for female penguins by species (Chinstrap, Gentoo, and Adelie)?

2. Do some exploring 

New functions: with + table for easy counting over multiple factors 

Update species name using case_when 

```{r}

counts <- with(penguins, table(species, sex)) 

penguins2 <- penguins %>% 
  mutate(common_name = 
           case_when(
            species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap", 
            species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo", 
            species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
           )
  ) %>% 
  select(-species) %>%  
  filter(sex == "FEMALE")
  
```

- Exploratory graphs 
- Summary statistics 
- Levene's tests for equal variances 
- One-Way ANOVA 
- Post-Hoc Testing 

```{r}
hists <- ggplot(penguins2, aes(x = flipper_length)) + 
  geom_histogram(aes(fill = common_name)) +  
  facet_wrap(~ comon_name)


qqs <- ggplot(penguins2, aes(sample = flipper_length)) + 
  geom_qq() + 
  facet_wrap(~ common_name)

# Levene's test for equal variances (> 2 groups... if you have two groups, use the F-test for equal variances)

# We'll use the function leveneTest from the 'car' package 

# H0: Varainces are equal 
# HA: Variances are unequal 

penguin_levene <- leveneTest(flipper_length ~ common_name, data = penguins2)
penguin_levene

#We retain the null hypothesis of eequal variances (p = 0.06)

variances <- penguins2 %>%  
  group_by(common_name) %>%  
  summarize(
    variance = var(flipper_length)
  )

variances

```

#3. One-Way ANOVA 
Single factor: common_name (species)
Number of levels in the factor: 3 (Chinstrap, Gentoo, and Adelie)
Random variable: flipper_length(mm)

Is there a significant difference in mean flipper length (mm) for female Chinstrap, Gentoo, and Adelie penguins? 

```{r}

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)

penguin_aov 
```




